On Extending a Full-Sharing Design with Batched Scheduling

### 2015-04-15 paper.tex
batched scheduling ->ok for now

In this paper we discuss a novel approach for solving the problem of
supporting multithreaded batched scheduling in the FS design and we
present a performance analysis comparison between local scheduling
with batched scheduling. Experimental results show that despite the
extra data structures required to combine FS design with batched
scheduling, the execution time of combination is still quite
competitive when comparing against the FS with local scheduling.

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=10.5cm]{figures/yap-mt.pdf}
\caption{Yap's multithreaded tabled space using the NS and FS designs}
\label{fig_yap_mt_support}
\end{figure}

@@

In this section, we show how we have extended the FS design to support
batched scheduling.

@@

Figure~\ref{fig_tabtries_pcc}(a) shows then a situation where two
threads, $T_1$ and $T_{k-1}$, are sharing the same subgoal entry call
$i$ when the subgoal is still under evaluation, i.e., the subgoal is
not yet complete. The current state of the evaluation shows an answer
trie with $3$ answers found for the subgoal call $i$. For the sake of
simplicity, we are omitting the internal answer trie nodes and we are
only showing the leaf nodes nodes $LN_1$, $LN_2$ and $LN_3$, in the
figure. With the PAC procedure, the leaf nodes are not chained in the
\emph{AT} data structure. Now, the chaining process is done privately,
and for that, we use the subgoal frame structure of each thread. On
the subgoal frame structure we added a new field, called
\emph{consumer}, to store the answers found within the execution of
the thread. In order to minimize the impact of the \emph{PAC}
optimization, each node within the new consumer answer structure has
two fields: (i) an entry pointer, which points to the corresponding
leaf node in the answer trie data structure; (ii) a next pointer to
chain the answers within the consumer structure. To maintain a good
performance, when the number of nodes exceeds a certain threshold, we
use a hash trie mechanism design similar to the one presented in the
work~\cite{Areias-ijpp15}. However, since this mechanism is private to
each thread, it does not require any of the tools that were necessary
to support concurrency. In particular, on each hash trie level, we
have removed the tools necessary to support concurrency, such as
useless pointers and compare-and-swap operations. We have chosen this
hashing mechanism, because it showed a good balance between lookup and
insert operations~\cite{Areias-ijpp15}, but the major reason was
mostly because of the integration in the TabMalloc memory
allocator~\cite{Areias-12b}.

@@

At the implementation level, the major difference between local and
batched scheduling is in the tabling operation \emph{tabled new
  answer}, where we decide what to do when an answer is found during
the evaluation. This operation checks whether a newly found answer is
already in the corresponding answer trie structure and, if not,
inserts it. The usage of batched scheduling with the FS design
requires further support since with batched scheduling, answers are
immediately propagated and we have to ensure that the propagation of
an answer occurs on all subgoal calls one and only once. To do so, we
take advantage of the private chaining procedure, presented in the
previous subsection, as a way to keep, for every subgoal call of every
thread, track of all the answers that were already propagated. This
requires minor changes to the \emph{tabled new answer} tabling
operation. Algorithm~\ref{alg_table_new_answer_batched} shows how we
have extended the tabled new answer operation to support the FS design
with batched scheduling.

>>>>

In this paper we discuss a novel approach for combining the FS design
with batched scheduling and we present a performance analysis
comparison between local scheduling with batched
scheduling. Experimental results show that despite the extra data
structures required to combine FS design with batched scheduling, the
execution time of combination is still quite competitive when
comparing against the FS with local scheduling.

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=9.5cm]{figures/yap-mt.pdf}
\caption{Yap's multithreaded tabled space using the NS and FS designs}
\label{fig_yap_mt_support}
\end{figure}

@@

In this section, we show how we have extended the FS design to support
batched scheduling. The usage of batched scheduling with the FS design
requires further support since with batched scheduling, answers are
immediately propagated and we have to ensure that the propagation of
an answer occurs on all subgoal calls one and only once. To do so,
next we show a procedure that, for every subgoal call of every thread,
keeps track of all the answers that were already propagated.

@@

Figure~\ref{fig_tabtries_pcc}(a) shows then a situation where two
threads, $T_1$ and $T_{k-1}$, are sharing the same subgoal entry call
$i$ when the subgoal is still under evaluation, i.e., the subgoal is
not yet complete. The current state of the evaluation shows an answer
trie with $3$ answers found for the subgoal call $i$. For the sake of
simplicity, we are omitting the internal answer trie nodes and we are
only showing the leaf nodes nodes $LN_1$, $LN_2$ and $LN_3$, in the
figure. With the PAC procedure, the leaf nodes are not chained in the
answer trie data structure. Now, the chaining process is done
privately, and for that, we use the subgoal frame structure of each
thread. On the subgoal frame structure we added a new field, called
\emph{consumer}, to store the answers found within the execution of
the thread. In order to minimize the impact of the \emph{PAC}
optimization, each node within the new consumer answer structure has
two fields: (i) an entry pointer, which points to the corresponding
leaf node in the answer trie data structure; (ii) a next pointer to
chain the answers within the consumer structure. To maintain a good
performance, when the number of nodes exceeds a certain threshold, we
use a hash trie mechanism design similar to the one presented in the
work~\cite{Areias-ijpp15}. However, since this mechanism is private to
each thread, it does not require any of the tools that were necessary
to support concurrency. In particular, on each hash trie level, we
have removed the tools necessary to support concurrency, such as
useless pointers and compare-and-swap operations. We have chosen this
hashing mechanism, because it showed a good balance between lookup and
insert operations~\cite{Areias-ijpp15}, but the major reason was
mostly because of the integration in the TabMalloc memory
allocator~\cite{Areias-12b}.

@@

At the implementation level, the major difference between local and
batched scheduling is in the tabling operation \emph{tabled new
  answer}, where we decide what to do when an answer is found during
the evaluation. This operation checks whether a newly found answer is
already in the corresponding answer trie structure and, if not,
inserts it. This requires minor changes to the \emph{tabled new
  answer} tabling
operation. Algorithm~\ref{alg_table_new_answer_batched} shows how we
have extended the tabled new answer operation to support the FS design
with batched scheduling.
